---
title: "HW6"
output: html_document
date: "2024-12-01"
---

```{r setup, include = FALSE}
library(tidyverse)
library(dplyr)
library(purrr)
library(ggplot2)
library(boot)
library(modelr)
library(mgcv)
options(scipen = 999) 
```

PROBLEM 2
```{r}
homicide = read_csv("homicide-data.csv")

#create city-state variable
homicide_df = homicide %>% drop_na() %>% mutate(city_state = str_c(city, state, sep = ", ")) 

#create binary variable indicating whether the homicide is solved
check_solved <- function(data, column, phrase) {
  data %>%
    mutate(solved = ifelse(grepl(phrase, .data[[column]]), "Yes", "No"))
}

homicide_df = check_solved(homicide_df, "disposition", "Closed by arrest")


#omit Dallas, TX; Phoenix, AZ; and Kansas City, MO and Tulsa, AL; 
#and making sure victim is Black or White
homicide_tidy = homicide_df %>% 
  filter(!(city_state == "Dallas, TX") ) %>% 
  filter(!(city_state == "Phoenix, AZ")) %>%  
  filter(!(city_state == "Kansas City, MO")) %>%  
  filter(!(city_state == "Tulsa, AL")) %>% 
  filter(victim_race == "Black" | victim_race == "White")


#victim_age
homicide_tidy = homicide_tidy %>% 
  mutate(victim_age = as.numeric(gsub("Unknown", "NA", victim_age)))



#creating binary variables
homicide_tidy = homicide_tidy %>%
  mutate(solved_binary = as.numeric(disposition == "Closed by arrest"),
         victim_race_binary = fct_relevel(victim_race, "White"),
         victim_sex_binary = fct_relevel(victim_sex, "Female"))

#glm for baltimore
homicide_baltimore = homicide_tidy %>% filter(city_state == "Baltimore, MD")
model1 = glm(solved_binary ~ victim_age + victim_race_binary + victim_sex_binary, data=homicide_baltimore, family=binomial())

model1_tidy = model1 %>% broom::tidy(conf.int=TRUE, conf.level=0.95, exponentiate=TRUE) %>%
  filter(term == "victim_sex_binaryMale") %>% 
  select(term, estimate, p.value, conf.low, conf.high) %>% 
  knitr::kable(digits = 3)
```
estimated odds ratio comparing male to female with all other variables fixed: 0.426, 95% CI: [0.324, 0.558]

```{r}
#glm for each city 

glm_city_results = homicide_tidy %>%
  nest(data = -city_state) %>% # nest data by city
  mutate(
    model2 = map(data, ~ glm(solved_binary ~ victim_age + victim_race_binary + victim_sex_binary, 
                            data = ., family = binomial())),
    model2_tidy = map(model2, ~ broom::tidy(., exponentiate = TRUE, conf.int = TRUE)) # Get ORs and CIs
  ) %>%
  unnest(model2_tidy) 


#Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot.
glm_city_results %>% 
  filter(term == "victim_sex_binaryMale") %>% 
  mutate(city_state = fct_reorder(city_state, estimate)) %>% 
  ggplot(aes(x = city_state, y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymax = conf.high, ymin = conf.low, width = 0.1)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(limits = c(0, 5)) +
  labs(x = "City", y = "Odds Ratio", title = "Odds Ratios Comparing Male to Female Victims in Solved Homicides (by City)")


```
Plot description: in the majority of cities, male victims are less likely to have their cases solved compared to female victims. New York has the lowest odds of a solved homicicde for male victims compared to female victims. New Mexico has the highest odds of a solved homicicde for male victims compared to female victims.


PROBLEM 3
```{r}
birthwt = read_csv("birthweight.csv")

#tidying data
#birthwt_tidy = birthwt %>% 
  drop_na() %>% mutate(
    fincome = fct_relevel(fincome, )
  )

#model 1
birthwt_model1 = lm(bwt ~ fincome + smoken + gaweeks, data=birthwt)

birthwt_model1_resid <- birthwt %>%
  modelr::add_predictions(birthwt_model1) %>%  
  modelr::add_residuals(birthwt_model1)

resid_pred_plot1 = birthwt_model1_resid %>% 
  ggplot(aes(x = pred, y = resid)) + 
  geom_point() +
  geom_smooth()


#model2
birthwt_model2 = lm(bwt ~ blength + gaweeks, data=birthwt)

#model3
birthwt_model3 = lm(bwt ~ bhead + blength + babysex + bhead*blength + blength*babysex + bhead*babysex + bhead*blength*babysex, data=birthwt)




#cross validation 
#creating testing and training
cv_df = crossv_mc(birthwt, 100) 

cv_df =
  cv_df %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))


#creating data frame with RMSE
cv_df_rmse = 
  cv_df %>% 
  mutate(
    linear_mod1 = map(train, ~birthwt_model1),
    linear_mod2 = map(train, ~birthwt_model2),
    linear_mod3 = map(train, ~birthwt_model3)
  ) %>% 
  mutate(
    rmse_mod1 = map2_dbl(linear_mod1, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_mod2 = map2_dbl(linear_mod2, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_mod3 = map2_dbl(linear_mod3, test, \(mod, df) rmse(model = mod, data = df))
  )

#plotting RMSE for each model
cv_df_rmse %>%  
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()

```
proposed regression model: financial income, average number of cigarettes smoked per day during pregnancy, gestational age in weeks as predictors. I picked these predictors because they are hypothesized factors that have an affect birthweight. Although I created this model based on hypothesized underlying factors, in the future I would check the Pearson correlation coefficients between the predictors and outcome before I used these variables as predictors. I decided to fit a linear model because all the variables are continuous.  

Model comparison: Model 3 has the lowest RMSE values and highest overall density, indicating it is the best fitted model. Model 1 has very high RMSE, meaning it is probably not well fitted (again, in the future I would check correlation coefficient before putting those variables into a model). Model 2 is better fitted than Model 1 and worst fitted than Model 3. 